Secure Composition \\ of Robust and Optimising Compilers

### 2024-05-27 csf2/main.tex
spec

To introduce the security argument of this paper, this section defines (security) properties, their satisfaction, and their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, this section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).

@@

This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}) and the unobservable event $\emptyevent$.
As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
\footnote{
Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
Prepending element $\event$ to a sequence $\trace$ uses the same notation: $\event\cdot\trace$.
}

Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
% 
The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
% 
The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
As an example, consider the trace:
$$\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$$
which describes the interaction with a memory where the deallocation of an address $\loc$ precedes a read (of some value $1729$) at that address in memory.
% 
This program behaviour is insecure w.r.t. a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.

In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
With this, we defined property satisfaction as follows:
\bul{a whole program $\wholeprogvar$ satisfies a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\pi$}.

@@

Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
Thus, from a security perspective, \Cref{def:propsat} considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, the we extend the concept of satisfaction with {\em robustness}.
Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of any features the programming language has, and the component is what is security-relevant.
%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
With this, \Thmref{def:propsat} can be extended as follows: for \bul{a component $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.

@@

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
This translation is considered {\em correct} if it is semantics-preserving.
That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ``compatible''.
Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
This is why {\em robust preservation} is a good candidate as a compiler-level security property~\cite{abate2019jour}.

This paper uses a general notion of robust preservation~\cite{abate2021extacc} that considers compilers that use languages with different trace models. 
To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
This relation induces the following two projection functions: (1) the existential image $\tau_\sim\left(\src{\pi}\right)$ and (2) the universal image $\sigma_\sim\left(\trg{\pi}\right)$.
These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
The case study of this paper exclusively uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:

@@

Using these projection functions, the more general version of robust preservation is defined as follows~\cite{abate2021extacc}.
% For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.

% \begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
%   % $\;$\\
%   % \vspace{-1em}
%   % \begin{nscenter}
%   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
%   \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
%   %, iff 
%   $\isdef$
%   \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
%   % \end{nscenter}
% \end{definition}
% 
A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\trg{\pi}$}.

\begin{definition}[Robust Preservation with $\sigma_\sim$]\label{def:rtp:sigma}
  % $\;$\\
  % \vspace{-1em}
  % \begin{nscenter}
  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
  \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\trg{\class}}{\sim}$}
  %, iff 
  $\isdef$
  \rul{$\forall \trg{\pi}\in\trg{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\sigma_\sim\left(\trg{\pi}\right)}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\trg{\pi}}$}.
  % \end{nscenter}
\end{definition}
Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\trg{\pi}}{\sim}$, even though $\trg{\pi}$ is a property and not a class.
A similar construction can be applied to the projection functions.\MP{wut?}

In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we have the classic definition of robust preservation~\cite{abate2019jour}.
\begin{definition}[Robust Preservation]\label{def:rtp}
  % $\;$\\
  % \vspace{-1em}
  % \begin{nscenter}
  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
  $\;$

@@

$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics. 
After branching or accessing memory, speculation can start and run subsequent code in a predetermined window until a rollback of operational state is performed. 
Within such windows, data is leaked, marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant, in this case just SPECTRE-PHT~\cite{kocher2019spectre}, whose semantics is sketched with \Cref{tr:e-ifz-true-spec}, where $\omega$ is the size of the speculation window.
Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
The semantics is kept general enough to allow for future extension to support different variants.

@@

\vspace{-.5em}
\[
  \begin{array}{rrcl}
    \text{(Base-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier}
  \end{array}
\]

@@

  %
  \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
    \ird{\cfstate[']}.\ird{\operatorname{win}} = \omega &
    \ird{\trace} = \ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag}
  }{
    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\begin{array}{c}\rtt{\ird{\cfstate['];\sandboxtag;0;\memstate}}{\ird{\expr[_1]}}\\\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr[_2]}}\end{array}}{\ird{\trace}}
  }{e-ifz-true-spec}
  %
  \typerule{$e-\ird{spec}-\text{eat}$}{
    \ird{\cfstate}.\ird{\operatorname{win}} > 0 &
    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr}}}{\rtt{\ird{\cfstate['];\sandboxtag;0;\memstate}}{\ird{\expr[']}}}{\ird{\event}}
    \\
    \ird{\cfstate['']}.\ird{\operatorname{win}} = \ird{\cfstate[']}.\ird{\operatorname{win}}-1
  }{
    \ghoststepto{\begin{array}{c}\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr}}\\\ird{r}\end{array}}{\begin{array}{c}\rtt{\ird{\cfstate[''];\sandboxtag;0;\memstate}}{\ird{\expr[']}}\\\ird{r}\end{array}}{\ird{\event}}
  }{e-spec-eat}
  % 
  \typerule{$e-\ird{spec}-\text{eaten}$}{
    \ird{r}.\ird{\cfstate}.\ird{\operatorname{win}} = 0
  }{
    \ghoststepto{\begin{array}{c}\ird{r}\\\ird{r'}\end{array}}{\ird{r'}}{\ird{Rlb}}
  }{e-spec-eaten}
  %
  \typerule{$e-\ird{spec}-\text{barrier}$}{
    \ird{r'}.\ird{\cfstate} = \ird{r}.\ird{\cfstate}[.\ird{\operatorname{win}} = 0]
  }{
    \ghoststepto{\begin{array}{c}\ird{r}\\\ird{r''}\end{array}}{\begin{array}{c}\ird{r'}\\\ird{r''}\end{array}}{\ird{Barrier}}
  }{e-spec-barrier}
\end{center}

The trace model is extended with $\ird{\ev{Spec}}$, $\ird{\ev{Rlb}}$, and $\ird{\ev{Barrier}}$ events that signal the respective operational action. 
A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode (see \Cref{tr:e-spec-barrier}) and does nothing when run in normal mode.
The speculative semantics are inspired by existing work~\cite{guarnieri2018spectector}.

>>>>

To introduce the security argument of this paper, this section first presents the concepts of (security) properties, of their satisfaction, and of their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, the section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).

@@

This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}) and the unobservable event $\emptyevent$.
As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
\footnote{
Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
Prepending elements to sequences uses the same notation: $\event\cdot\trace$.
}

Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances.
As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
This program behaviour is insecure w.r.t. a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.

In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
With this, property satisfaction is defined as follows:
\bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).

@@

Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, the concept of satisfaction can be extended with {\em robustness}.
Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has, and the component is what is security-relevant.
%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.

@@

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
This translation is considered {\em correct} if it is semantics-preserving.
That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ''compatible''.
Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
This is why {\em robust preservation} is a good candidate as a compiler-level security property~\cite{abate2019jour}.

This paper uses a more general notion of robust preservation~\cite{abate2021extacc} that applies to compilers that use languages with different trace models. 
To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
Using this relation, there are two projection functions: (1) the existential image $\tau_\sim\left(\src{\pi}\right)$ and (2) the universal image $\sigma_\sim\left(\trg{\pi}\right)$.
These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties. 
The case study of this paper exclusively uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:

@@

Using these projection functions, the more general version of robust preservation is defined as follows~\cite{abate2021extacc}:
% For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.

% \begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
%   % $\;$\\
%   % \vspace{-1em}
%   % \begin{nscenter}
%   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
%   \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
%   %, iff 
%   $\isdef$
%   \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
%   % \end{nscenter}
% \end{definition}
% 
A \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\trg{\pi}$}.

\begin{definition}[Robust Preservation with $\sigma_\sim$]\label{def:rtp:sigma}
  % $\;$\\
  % \vspace{-1em}
  % \begin{nscenter}
  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
  \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\trg{\class}}{\sim}$}
  %, iff 
  $\isdef$
  \rul{$\forall \trg{\pi}\in\trg{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\sigma_\sim\left(\trg{\pi}\right)}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\trg{\pi}}$}.
  % \end{nscenter}
\end{definition}
Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\trg{\pi}}{\sim}$, even though $\trg{\pi}$ is a property and not a class.
A similar construction can be applied to the projection functions.

A degenerate case where $\sim$ coincides with equality, i.e., the trace model is the same for both source and target programs and the compiler does not have an effect on the resulting traces, is the classic definition of robust preservation~\cite{abate2019jour}:

\begin{definition}[Robust Preservation]\label{def:rtp}
  % $\;$\\
  % \vspace{-1em}
  % \begin{nscenter}
  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
  $\;$

@@

$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics. 
After branching or accessing memory, speculation can start, pushes the current configuration on the speculation register ($\ird{S}$), and run subsequent code in a predetermined window until a rollback of operational state is performed. 
The window is set inside the control-flow state ($\ird{\cfstate}$), whose details are in the technical report.
Within such windows, data is leaked, marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant, in this case just SPECTRE-PHT~\cite{kocher2019spectre}, whose semantics is sketched with \Cref{tr:e-ifz-true-spec}, where $\omega$ is the size of the speculation window.
Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
The semantics is kept general enough to allow for future extension to support different variants.

@@

\vspace{-.5em}
\[
  \begin{array}{rrcl}
    \text{(Base-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier} \\
    \text{(Speculation Register)} & \ird{S} &:=& \ird{\hole{\cdot}} \mid \left(\ird{\Omega};\ird{\expr}\right)
  \end{array}
\]

@@

  %
  \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
    \ird{\cfstate[']}.\ird{\operatorname{win}} = \omega &
    \ird{\trace} = \ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag} \\
    \ird{S} = ({\ird{\cfstate;\sandboxtag;D;\memstate}};{\ird{\expr[_2]}}),\ird{\hole{\cdot}}
  }{
    \pstepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate});\ird{\hole{\cdot}}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate});\ird{S}}{\ird{\expr[_1]}}}{\ird{\trace}}
  }{e-ifz-true-spec}
  %
  \typerule{$e-\ird{spec}-\text{eat}$}{
    \ird{\cfstate}.\ird{\operatorname{win}} > 0 &
    \pstepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\expr}}}{\rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\expr[']}}}{\ird{\event}}
    \\
    \ird{\cfstate['']}.\ird{\operatorname{win}} = \ird{\cfstate[']}.\ird{\operatorname{win}}-1 &
    \ird{S} \not=\ird{\hole{\cdot}}
  }{
    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr}}}{\rtt{(\ird{\cfstate[''];\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr[']}}}{\ird{\event}}
  }{e-spec-eat}
  % 
  \typerule{$e-\ird{spec}-\text{eaten}$}{
    \ird{\cfstate}.\ird{\operatorname{win}} = 0 &
    \ird{S} = \left(\ird{\Omega};\ird{\expr[']}\right)
  }{
    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr}}}{\rtt{(\ird{\Omega};\ird{\hole{\cdot}})}{\ird{\expr[']}}}{\ird{\event}}
  }{e-spec-eaten}
  %
  \typerule{$e-\ird{spec}-\text{barrier}$}{
    \ird{\cfstate'} = \ird{\cfstate}[.\ird{\operatorname{win}} = 0] &
    \ird{S} \not=\ird{\hole{\cdot}}
  }{
    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\lbarrier}}}{\rtt{(\ird{\cfstate';\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\lbarrier}}}{\ird{\event}}
  }{e-spec-barrier}
\end{center}

The trace model is extended with $\ird{\ev{Spec}}$, $\ird{\ev{Rlb}}$, and $\ird{\ev{Barrier}}$ events that signal the respective operational action. 
A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.
\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which should be set to $0$.}

The speculative semantics are inspired by existing work~\cite{guarnieri2018spectector}.
